Equilibria and Convergence of Fictitious Play on Network Aggregative Games

### 2021-05-25 Documents/FP in NAGs/neurips_2021.tex
Within Page Limit

In this study, we investigate a variant of the aggregative game known as the \emph{Network
Aggregative} (NA) Game. This format assumes that there is an underlying communication network
through which agents interact. Then each agent updates their actions according only to those agents
with whom they communicate. This significantly relaxes the communication load on each agent and
lifts the need for a central coordinator. Recent work on the NA game has shown that it is possible
for agents to reach an equilibrium strategy in an entirely distributed manner \cite{Koshal2016, Shokri2020, Shokri2021, Parise2015}. We contribute in this direction by analysing the long term behaviour of learning on an NA game. In particular, we analyse the Fictitious Play Learning
Algorithm \cite{Brown1949, Harris1998}, in which agents are assumed to be myopic, in that they react solely
to the previous behaviour of the others.

@@

  Using this, we follow in the footsteps of Ewerhart \cite{Ostrovski2014} and Harris \cite{Harris1998} to define
  Fictitious Play in continuous time, but with a slight adaptation for the NA game.
%
  \begin{definition}[Network Aggregative - Continuous Time Fictitious Play (NA-CTFP)] \label{def::NACTFP}
    An NA-CTFP is defined as a measurable map $m$ with components $m^\mu$ such that for all $\mu$
    and all $t \geq 1$, $m^\mu: [0, \infty) \rightarrow \Delta_\mu$ satisfies $m^\mu(t) \in
    BR^\mu(\alpha_{\sigma}^\mu)$ for almost all $t \geq 1$.
  \end{definition}

@@

  \begin{assumption}\label{ass::matrixgame}
    The payoffs are given through matrix games and, therefore, are bilinear. Payoff matrices
    have a rich history in game theory and has generated a number of prototypical examples for
    economic study including the Prisoner's Dilemma game (see \cite{Axelrod1980} for an interesting
    implementation of this). They also allow for the design of multi-agent systems in
    computational settings, particularly in the case of task and resource allocation \cite{Nisan2007}. It should be noted, however, that game
    theoretic analysis is starting to consider various other forms of utility functions,
    including monotone \cite{Tatarenko2020} and convex \cite{Parise2020}. We believe that the analysis of
    Fictitious Play should follow in these developments and we consider it as an important area
    of future work.
  \end{assumption}

@@

  \begin{assumption}\label{ass::zerosum}
    The NA game is zero-sum in the sense that $\sum_{\mu} u^\mu(\xmu, \weightedsum)$ for any set
    of states $(x^\mu)_{\mu \in N^\mu}$. This is, perhaps, one of the stronger assumptions in
    our analysis, and is required for the fixed point analysis. However, in Section
    \ref{sec::CCEConvergence}, we perform a regret analysis that considers the long term
    behaviour of NA-CTFP without this assumption.
  \end{assumption}

@@

  Note
  that, in order to get this formulation, we had to use the assumption
  of payoffs being bilinear so that we could separate out the term in
  the weighted sum involving $x$ from $\bar{x}^\nu$. In particular, we used this assumption to move from (\ref{eq::bilinear1}) to (\ref{eq::bilinear2}).

@@

With these results in place, we can show that NA-CTFP converges
to a fixed point. In particular, let $\Omega(\alpha)$ be the set of
all limit points for $\alpha(t)$. Then, a NA-CTFP path is said to have
converged if $\Omega(\alpha)$ is contained within the set of Nash
Equilibria of the game. We adapt the techniques
of \cite{Ewerhart2020} to prove that zero-sum NA games have the CTFP
property.
%
  \begin{theorem} \label{thm::NACTFPtoFixed}
    Any zero-sum NA game has the property that, for any NA-CTFP path $m$, $\alpha(t; m)$ (i.e. the time-averaged state) converges to a set of fixed points.
  \end{theorem}

@@

  For an NA game, if a set of actions $s = (s^1, \ldots, s^N)$ is
  drawn from a joint probability distribution $\mathcal{D}$, then
  there is a corresponding set of reference vectors $\sigma =
  (\sigma^1, \ldots, \sigma^N)$, where $\sigma^\mu = \sum_{\nu \in
    N^\mu} \wmunu s^\nu$.  Therefore, $\mathcal{D}$ is a probability
  distribution over all actions and all references.  Consider, then, a
  vector of joint distributions over actions and references which is
  generated by NA-CTFP. In particular, if by playing with NA-CTFP, the
  agents reach the state $(\xmu)_{i = 1}^N$ with references
  $(\refmu)_{i = 1}^N$, this is the vector $\mathcal{D} =
  (\mathcal{D}^1, ..., \mathcal{D}^N)$ such that
  $(\mathcal{D}^\mu)_{ij} = \xmu_i \refmu_j$. Then, the expected
  payoff that the agent would receive for playing this strategy is
%
  \begin{align*}
    u^\mu(\xmu, \refmu) & = \xmu \cdot A^\mu \refmu = \sum_{i, j} (A^\mu)_{ij} \xmu_i \refmu_j \nonumber 
  \end{align*}

@@

  \begin{remark} \label{rem::regretvsCCE}
    The notion of the CCE set is related to the idea of \emph{average external regret} (for the
    sake of brevity we drop the term `external' and refer the reader to \cite{Nisan2007}). Here, we
    will present what is meant by average regret and state that if at some time $T$ all agents'
    average regret is non-positive, then the game is said to have reached the CCE set. The
    reader should consult \cite{Ostrovski2014} for an excellent exposition regarding the
    link between the CCE set and average regret in two player games which, of course, extends
    naturally to the NA game.

@@

  \begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = \textwidth]{Figures/convergentShapley.png}
    \caption{\label{fig::convergentShapley}}
    \end{subfigure}
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = \textwidth]{Figures/nonConvergentShapley.png}
      \caption{\label{fig::nonconvergentShapley}}
    \end{subfigure}
    \caption{\label{fig::Shapley} Orbits of the Fictitious Play in the Three Player Chain (c.f.
    Figure \ref{fig::ThreePlayerNetwork}) with payoff matrices given by (a)
    (\ref{eq::zeroSumShapley}) showing convergence to the interior NE (b)
    (\ref{eq::nonzeroSumShapley}) showing cycles around the interior NE.}
  \end{figure}

>>>>

In this study, we investigate a variant of the aggregative game known as the \emph{Network
Aggregative} (NA) Game. This format assumes that each agent updates their actions according only to those agents with whom they are connected on an underlying network. This significantly relaxes the communication load on each agent and
lifts the need for a central coordinator. Recent work on the NA game has shown that it is possible
for agents to reach an equilibrium strategy in an entirely distributed manner \cite{Koshal2016, Shokri2020, Shokri2021, Parise2015}. We contribute in this direction by analysing the long term behaviour of learning on an NA game. In particular, we analyse the Fictitious Play Learning
Algorithm \cite{Brown1949, Harris1998}, in which agents are assumed to be myopic, in that they react solely to the previous behaviour of the others.

@@

  Using this, we follow in the footsteps of  \cite{Ewerhart2020} and \cite{Harris1998} to define
  Fictitious Play in continuous time, but with a slight adaptation for the NA game.
%
  \begin{definition}[Network Aggregative - Continuous Time Fictitious Play (NA-CTFP)] \label{def::NACTFP}
    An NA-CTFP is defined as a measurable map $m$ with components $m^\mu$ such that for all $\mu$
    and all $t \geq 1$, $m^\mu: [0, \infty) \rightarrow \Delta_\mu$ satisfies $m^\mu(t) \in
    BR^\mu(\alpha_{\sigma}^\mu)$ for almost all $t \geq 1$.
  \end{definition}

@@

  \begin{assumption}\label{ass::matrixgame}
    The payoffs are given through matrix games and, therefore, are bilinear. Payoff matrices
    have a rich history in game theory and allow for the design of multi-agent systems in
    computational settings, particularly in the case of task and resource allocation \cite{Nisan2007}. It should be noted, however, that game
    theoretic analysis is starting to consider various other forms of utility functions,
    including monotone \cite{Tatarenko2020} and convex \cite{Parise2020}. We believe that the analysis of
    Fictitious Play should follow in these developments and we consider it as an important area
    of future work.
  \end{assumption}

@@

  \begin{assumption}\label{ass::zerosum}
    The NA game is zero-sum in the sense that $\sum_{\mu} u^\mu(\xmu, \weightedsum) = 0$ for any set
    of states $(x^\mu)_{\mu \in N^\mu}$. This is, perhaps, one of the stronger assumptions in
    our analysis, and is required for the fixed point analysis. However, in Section
    \ref{sec::CCEConvergence}, we perform a regret analysis that considers the long term
    behaviour of NA-CTFP without this assumption.
  \end{assumption}

@@

  Note that, in order to get this formulation, we had to use Assumption \ref{ass::matrixgame}
 to move from (\ref{eq::bilinear1}) to (\ref{eq::bilinear2}).

@@

With these results in place, we can show that NA-CTFP converges
to a fixed point. In particular, let $\Omega(\alpha)$ be the set of
all limit points for $\alpha(t)$. Then, a NA-CTFP path is said to have
converged if $\Omega(\alpha)$ is contained within the set of Nash
Equilibria of the game. We adapt the techniques
of \cite{Ewerhart2020} to prove that this is the case.
%
  \begin{theorem} \label{thm::NACTFPtoFixed}
    Any zero-sum NA game has the property that, for any NA-CTFP path $m$, $\alpha(t; m)$ (i.e. the time-averaged state) converges to a set of fixed points.
  \end{theorem}

@@

  For an NA game, if a set of actions $s = (s^1, \ldots, s^N)$ is
  drawn from a joint probability distribution $\mathcal{D}$, then
  there is a corresponding set of reference vectors $\sigma =
  (\sigma^1, \ldots, \sigma^N)$, where $\sigma^\mu = \sum_{\nu \in
    N^\mu} \wmunu s^\nu$.  Therefore, $\mathcal{D}$ is a probability
  distribution over both the actions and the references.  Consider, then, a
  vector of joint distributions over actions and references which is
  generated by NA-CTFP. In particular, if by playing with NA-CTFP, the
  agents reach state $(\xmu)_{i = 1}^N$ with references
  $(\refmu)_{i = 1}^N$, $\mathcal{D} =
  (\mathcal{D}^1, ..., \mathcal{D}^N)$ is such that
  $(\mathcal{D}^\mu)_{ij} = \xmu_i \refmu_j$. Then, the expected
  payoff that the agent would receive for playing this strategy is
%
  \begin{align*}
    u^\mu(\xmu, \refmu) & = \xmu \cdot A^\mu \refmu = \sum_{i, j} (A^\mu)_{ij} \xmu_i \refmu_j \nonumber 
  \end{align*}

@@

  \begin{remark} \label{rem::regretvsCCE}
    The notion of the CCE set is related to the idea of \emph{average regret}  \cite{Nisan2007}. Here, we
    will present what is meant by average regret and state that if at some time $T$ all agents'
    average regret is non-positive, then the game is said to have reached the CCE set. The
    reader should consult \cite{Ostrovski2014} for an excellent exposition regarding the
    link between the CCE set and average regret in two player games which, of course, extends
    naturally to the NA game.

@@

  \begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = 0.8\textwidth]{Figures/convergentShapley.png}
    \caption{\label{fig::convergentShapley}}
    \end{subfigure}
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = 0.8 \textwidth]{Figures/nonConvergentShapley.png}
      \caption{\label{fig::nonconvergentShapley}}
    \end{subfigure}
    \caption{\label{fig::Shapley} Orbits of the Fictitious Play in the Three Player Chain (c.f.
    Figure \ref{fig::ThreePlayerNetwork}) with payoff matrices given by (a)
    (\ref{eq::zeroSumShapley}) showing convergence to the interior NE (b)
    (\ref{eq::nonzeroSumShapley}) showing cycles around the interior NE.}
  \end{figure}
